Influences of GDP in the U.S

### 2022-11-14 Manuscript/manuscript.tex
commit

\citet{divya2014study} found that exchange rates and market indexes are important factors that influence an economies GDP. 
They also found that inflation is highly correlated, but not a significant influencer.  
\citet{van2009gdp} researches the critism behind GDP's influence in economies. They compare the understanding that even though GDP can influence economically relevant decisions, it does not factor in social welfare. 
Since GDP is a global indicator "The importance of GDP information for firms, investors and citizens/consumers is illustrated by the media – television, radio, newspapers, financial and other magazines, and internet – informing us on a daily basis about the status of our national GDP, both over time and in comparison with other countries" (\citet[]{van2009gdp}).
They also touch on how government agencies and politicians strive to avoid low GDP. When GDP is low, this can lead to negative voter response, and less public expenditures, which are both fatal prospects for those in power (\citet[]{van2009gdp}).
\citet{szustak2021relationship} found that the relationship between power production and GDP is random. This 
\citet{kalyoncu2013causality} found that there is "unidirectional causality from per capita GDP to per capita energy consumption for Armenia." Their research focused on the relationship between energy consumption and economic growth.
The countries they focused their study on, Georgia, Azerbaijan and Armenia from 1995 to 2009, all faced low energy supply.

For this paper, we will quantify the relation between influencers of GDP. 
The main focus is to test whether energy consumption has a significant relationship to Gross Domestic Product. 
Since energy data is readily avaiable on a monthly basis, this could be a factor economists use to predict GDP on a monthly scale.

>>>>

\citet{divya2014study} found that Exchange rate, Sensex and Balance of Payment reflected by current and capital account balance are important factors that influence India’s GDP. 
Using correlation and ANOVA, they also found that inflation is highly correlated, but not a significant influencer. 
The main purpose of their study is to identify key signals for an economy before a crisis occurs.

\citet{van2009gdp} researches the criticism behind GDP's influence in economies. 
They compare the understanding that even though GDP can influence economically relevant decisions, it does not factor in social welfare.
Since GDP is a global indicator 'The importance of GDP information for firms, investors and citizens/consumers is illustrated by the media – television, radio, newspapers, financial and other magazines, and internet – informing us on a daily basis about the status of our national GDP, both over time and in comparison with other countries.'(\citet[]{van2009gdp})
They also touch on how government agencies and politicians strive to avoid low GDP. When GDP is low, this can lead to negative voter response, and less public expenditures, which are both fatal prospects for those in power(\citet[]{van2009gdp}).

\citet{szustak2021relationship} investigated the possible relationship between energy production and GDP growth. 
Their goal was to discover the direction of the relationship if one exists. 
found that the relationship between power production and GDP is random. 
The study found that there is not a relationship between energy production and GDP. 
Even though in some counties the relationship was stronger, they concluded that the relationship was random.

\citet{kalyoncu2013causality} focused on the relationship between energy consumption and economic growth in Georgia, Azerbaijan and Armenia from 1995 to 2009. 
Using the Engle-Granger cointegration and Granger causality tests, they investigated the direction of causality for the purpose of informing policy makers. 
Their research found that there is `unidirectional causality from per capita GDP to per capita energy consumption for Armenia.' (\citet[]{kalyoncu2013causality}) 
The countries they focused their study on all faced low energy supply which is a factor that must be taken into account.

This paper strives to quantify the relation between influencers of GDP. 
The main focus is to test whether energy consumption has a significant relationship to Gross Domestic Product. 
This will be conducted using the variable Total Primary Energy Consumption. 
Since energy data is readily available on a monthly basis, this indicator could be used by economists to predict GDP on a monthly scale. 
The goal is to be able to identify signals of crucial moments to allow policy makers to make more informed decisions.
